First Exclusive Measurement of Deep Virtual Compton Scattering off $^4$He: Toward the 3D tomography of nuclei

### 2017-01-24 eg6-letters/DVCS-coh-He4/dvcs_he4_prl.tex
added the theoretical models, the results interpretations and the summary

A DVCS process on a nuclear target differ from single nucleon scattering in 
providing access to the measure two DVCS channels. In the coherent DVCS 
channel, the target nucleus ($A$) remains intact and recoils as a whole while 
emitting a real photon ($eA \rightarrow e' A' \gamma$), allowing to measure the 
nuclear GPDs of the target. In the incoherent channel, the nucleus breaks up 
and the DVCS takes place on a bound nucleon ($N$) that emits the final photon 
($eA \rightarrow e' N' \gamma$ X), enabling the GPDs measurement of the bound 
nucleons and to study their modifications in the nuclear medium via the GPDs.  
Figure \ref{fig:diags} illustrates the dominant mechanism for the coherent DVCS 
channel on $^4$He. At sufficiently large squared electron momentum transfer 
$Q^2$ ($= -(k-k')^{2}$) and small squared momentum transfer $t$ ($= 
(p-p')^{2}$), the QCD factorization theorem predicts that the DVCS handbag 
diagram can be factorized into two parts, hard and soft parts 
\cite{Freund_Collins,Ji_Osborne}. The hard part includes photons-quark 
interaction and it is calculable through perturbative methods, while the 
soft/non-perturbative part is parametrized in terms of GPDs, which embed the 
partonic structure of the hadron.  
\begin{figure}[tb]
\includegraphics[width=6.5cm]{figs/DVCS_diagram.pdf}
\vspace{-0.cm}
\caption{Deep virtual Compton scattering process in the handbag approximation.}
\label{fig:diags}
\end{figure}

@@

%DVCS selection
Identifying the coherent DVCS candidates is the first step of the data 
analysis. These events have one electron, one $^4$He, and at least one photon 
in the final state. Electrons were identified by passing the fiducial cuts and 
having signals in all the sub-detectors of CLAS spectrometer (drift chambers, 
Cherenkov counters, the standard CLAS electromagnetic calorimeter, and 
scintillators). $^4$He tracks were identified by passing all the geometrical, 
timing and quality cuts in the RTPC detector. The most energetic IC photon was 
considered as the DVCS photon candidate. Next, a $Q^{2}>1~[GeV^{2}/c^{2}]$ cut 
is applied on the DVCS candidates in order to ensure that the interaction 
occurs at the partonic level and the applicability of the factorization in the 
DVCS handbag diagram. Once the three final state particles were identified with 
their 3-momentum vectors, the exclusivity of the coherent DVCS events were 
ensured by applying a set of exclusive cuts, which are: the co-planarity angle 
($\Delta \phi$), missing energy, missing mass squared and missing transverse 
momentum in the $e'^4He'\gamma X$ final state configuration, the missing mass 
squared in the $e'^4He'X$ and $e'\gamma X$ configurations, and finally the cone 
angle ($\theta$) between the measured real photon and the missing particle in 
the $e'^4He'X$ configuration. Figure~\ref{fig:kin-cuts} presents four of the 
applied exclusivity cuts, where 3$\sigma$ cuts are applied on all the exclusive 
quantities except the missing energy, for which a [-0.45,0.5] GeV cut was 
adopted to reduce the background contribution.

@@

Figure \ref{fig:A_LU-coh} presents the coherent $A_{LU}$ for the three
sets of two-dimensional bins. The asymmetries are fitted with the form of
equation \ref{eq:A_LU-coh}, where the real and the imaginary part of the CFF
$\mathcal{H}_{A}$ are the free parameters in the fit. Figure \ref{fig:alu90} 
shows the $Q^2$, $x_{B}$, and $-t$-dependencies of the fitted $A_{LU}$ signals 
at $\phi$~=~90$^{\circ}$. The $x_{B}$ and $-t$-dependencies are compared to 
theoretical calculations performed by S.~Liuti and K.~Taneja 
\cite{simonetta_2}.  Their model relies on the impulse approximation and uses 
advanced spectral function of the nuclei to calculate the nuclear GPDs and then 
the observables. The calculations were carried out at slightly different 
kinematics than ours but provide already some guidance. The experimental 
results appear to have larger asymmetries compared to the calculations.  These 
differences may arise from nuclear effects which are not taken into account in 
the model, such as long-range interactions. Our measurements also agree with 
those of HERMES, considering their large uncertainties.

@@

\begin{figure}[tb]
\includegraphics[width=8.9cm]{figs/Coherent_CFF.pdf}
\caption{The model-independent extraction of the imaginary (blue points) and
real (red points) parts of the $^4$He CFF $\mathcal{H}_A$, as functions of
$Q^{2}$ (on the top right), $x_B$ (on the top left), and $t$ (on the bottom).  
The full red curves are calculations based on the impulse approximation from
\cite{Vadim_priv}. The black-dashed curves in the right plots are calculations 
from a convolution model based on the VGG model for the nucleons' GPDs 
\cite{Guidal_priv}. The blue long-dashed curve on the top-right plot is from
the off-shell calculations based on reference \cite{GonzalezHernandez:2012jv}.}
\label{fig:CFF_HA}
\end{figure}

%interpretations

%conclusion

@@

We thank the staff of the Accelerator and Physics Divisions
at Jefferson Lab for making this experiment possible.

@@

@@

@@

>>>>

A DVCS process on a nuclear target differ from single nucleon scattering in 
providing access to the measure two DVCS channels. In the coherent DVCS 
channel, the target nucleus ($A$) remains intact and recoils as a whole while 
emitting a real photon ($eA \rightarrow e' A' \gamma$), allowing to measure the 
nuclear GPDs of the target. In the incoherent channel, the nucleus breaks up 
and the DVCS takes place on a bound nucleon ($N$) that emits the final photon 
($eA \rightarrow e' N' \gamma$ X), enabling the GPDs measurement of the bound 
nucleons and to study their modifications in the nuclear medium via the GPDs.  
Figure \ref{fig:diags} illustrates the dominant mechanism for the coherent DVCS 
channel on $^4$He. At sufficiently large squared electron momentum transfer 
$Q^2$ ($= -(k-k')^{2}$) and small squared momentum transfer $t$ ($= 
(p-p')^{2}$), the QCD factorization theorem predicts that the DVCS handbag 
diagram can be factorized into two parts, hard and soft parts 
\cite{Freund_Collins,Ji_Osborne}. The hard part includes photons-quark 
interaction and it is calculable through perturbative methods, while the 
soft/non-perturbative part is parametrized in terms of GPDs, which embed the 
partonic structure of the hadron.  
\begin{figure}[tb]
\includegraphics[width=6.5cm]{figs/DVCS_diagram.pdf}
\caption{Representation of the leading-order, twist-2, handbag diagram of the 
DVCS process off $^4$He.}
\label{fig:diags}
\end{figure}

@@

%DVCS selection
Identifying the coherent DVCS candidates is the first step of the data 
analysis. These events have one electron, one $^4$He, and at least one photon 
in the final state. Electrons were identified by passing the fiducial cuts and 
having signals in all the sub-detectors of CLAS spectrometer (drift chambers, 
Cherenkov counters, the standard CLAS electromagnetic calorimeter, and 
scintillators). $^4$He tracks were identified by passing all the geometrical, 
timing and quality cuts in the RTPC detector. In principle, with constant beam 
luminosity, target density and pressure, the event rate has to be constant over 
the experimental time. Due to the changes in the experimental conditions, such 
as changing a trigger in a detector, a slight shift in the beam position or a 
system failure somewhere, this rate changes. We minimize the effects of these 
changes on the reconstructed events by selecting the good runs. To this aim, we 
monitor the ratio between the number of the good tracks reconstructed in the 
RTPC to number of the detected good electrons in the CLAS detector as a 
function of run number \cite{eg6_note}.

Even though the DVCS reaction has only one real photon in the final state, 
events with more than one good photon are not discarded at this stage. This is 
motivated by the fact that some photons correspond to random coincidences and 
discarding these events results in losing good events. Then, events with one or 
more $\pi^{0}$ are removed from the coherent DVCS sample. After that, the most 
energetic IC photon was considered as the DVCS photon candidate. Next, a 
$Q^{2}>1~[GeV^{2}/c^{2}]$ cut is applied on the DVCS candidates in order to 
ensure that the interaction occurs at the partonic level and the applicability 
of the factorization in the DVCS handbag diagram.  Once the three final state 
particles were identified with their 3-momentum vectors, the exclusivity of the 
coherent DVCS events were ensured by applying a set of exclusive cuts, which 
are: the co-planarity angle ($\Delta \phi$), missing energy, missing mass 
squared and missing transverse momentum in the $e'^4He'\gamma X$ final state 
configuration, the missing mass squared in the $e'^4He'X$ and $e'\gamma X$ 
configurations, and finally the cone angle ($\theta$) between the measured real 
photon and the missing particle in the $e'^4He'X$ configuration.  
Figure~\ref{fig:kin-cuts} presents four of the applied exclusivity cuts, where 
3$\sigma$ cuts are applied on all the exclusive quantities except the missing 
energy, for which a [-0.45,0.5] GeV cut was adopted to reduce the background 
contribution. Finally,

@@

Figure \ref{fig:alu} presents the coherent $A_{LU}$ for the three
sets of two-dimensional bins. The asymmetries are fitted with the form of
equation \ref{eq:A_LU-coh}, where the real and the imaginary part of the CFF
$\mathcal{H}_{A}$ are the free parameters in the fit. Figure \ref{fig:alu90} 
shows the $Q^2$, $x_{B}$, and $-t$-dependencies of the fitted $A_{LU}$ signals 
at $\phi$~=~90$^{\circ}$. The $x_{B}$ and $-t$-dependencies are compared to 
theoretical calculations performed by S.~Liuti and K.~Taneja 
\cite{simonetta_2}. Their model relies on the impulse approximation and uses 
advanced spectral function of the nuclei to calculate the nuclear GPDs and then 
the observables. The calculations were carried out at slightly different 
kinematics than ours but provide already some guidance. The experimental 
results appear to have larger asymmetries compared to the calculations.  These 
differences may arise from nuclear effects which are not taken into account in 
the model, such as long-range interactions. Our measurements also agree with 
those of HERMES, considering their large uncertainties.

@@

As has been advertised previously, the $^4$He CFF $\mathcal{H}_A$ can be 
extracted from fitting the experimentally measured coherent $A_{LU}$ in a 
totally model independent way, which is not the case of nucleon targets. For 
the later, four CFFs exist and extracting them is always made by making 
limitations according some theoretical models and by neglecting some CFFs 
\cite{Jo:2015ema}. Figure \ref{fig:CFF_HA} presents our extracted imaginary and 
real parts of the CFF $\mathcal{H}_A$ as function of the kinematical variable 
($Q^{2}$, $x_B$,$-t$), and compared to some theoretical calculations.

\begin{figure}[tb]
\includegraphics[width=8.9cm]{figs/Coherent_CFF.pdf}
\caption{The model-independent extraction of the imaginary (top panel) and
real (bottom panel) parts of the $^4$He CFF $\mathcal{H}_A$, as functions of
$Q^{2}$ (right panel), $x_B$ (middle panel), and $t$ (left panel). The full red 
curves are calculations based on an on-shell model from
\cite{Vadim_priv}. The black-dashed curves are calculations from a convolution 
model based on the VGG model for the nucleons' GPDs \cite{Guidal_priv}. The 
blue long-dashed curve on the top-right plot is from
an off-shell model based on \cite{GonzalezHernandez:2012jv}.}
\label{fig:CFF_HA}
\end{figure}

We display in figure \ref{fig:CFF_HA} calculations for $\mathcal{H}_A$ from 
three GPD models: On-shell, Convolution, and off-shell models. In the On-shell 
model \cite{Vadim_priv}, a nucleus is assumed to be composed of 
non-relativistic non-interacting nucleons, and these nucleons interact 
independently with the probe. For the nucleons, the GPDs are modeled according 
to the dual parametrization \cite{Guzey:2006xi}. In the Convolution model, the 
same assumption has been made for the nucleus, while the nucleon GPDs were 
extracted from the VGG model, which is based on the double distributions ansatz 
\cite{DD_model}. The Off-shell model \cite{GonzalezHernandez:2012jv} relies on 
the impulse approximation also, but uses advanced spectral function of the 
nuclei that accounts for all configurations of the final nuclear system and the 
binding effects between the nucleons.

In figure \ref{fig:CFF_HA}, within the given uncertainties, the extracted CFF 
shows a slight dependence on $Q^{2}$, $x_B$, and $-t$, which are in agreement 
with the theoretical calculations. One can see a difference between the 
precision of the extracted imaginary and real parts, which is expected because 
$\alpha_2$ is suppressed compared to $\alpha_0$ contribution. However, we note 
that the error bars are finite and that the fit converge without placing any 
bound on the CFF, which is necessary on proton targets.

%conclusion
In summary, we presented the first exclusive measurement of the coherent DVCS 
off $^4$He using CLAS spectrometer, supplemented with an inner calorimeter, a 5 
Tesla solenoid, and a specially designed radial TPC. This dataset represents a 
unique source for the nuclear DVCS global dataset, which will be used to 
constrain GPD models. The measured beam-spin asymmetries show a very strong 
signal and allowed to perform the first fully model-independent extraction of 
the $^4$He CFF $\mathcal{H}_A$. The extracted CFFs, while limited by 
statistics, are in a good agreement with the available GPD models. This opens 
many new perspectives to study the nuclear structure within the GPDs framework 
and pave the way for future more precise measurements at JLab 12~GeV program 
and possibly at the Electron-Ion Collider (EIC) to achieve better understanding 
of the nuclear effects.

@@

We acknowledge the staff of the Accelerator and Physics Divisions at Jefferson 
Lab for making this experiment possible. This work is supported by the U.S.  
Department of Energy, Office of Science, Office of Nuclear Physics contract 
DE-AC05-06OR23177.

@@

\bibitem{eg6_note}
M. Hattawy {\it et al.} (CLAS-EG6 Working Group), CLAS internal analysis note, 
2016.

@@

\bibitem{Guzey:2006xi} V.~Guzey and T.~Teckentrup,
     Phys.\ Rev.\ D {\bf 74}, 054027 (2006)

@@

\bibitem{DD_model}
I.~V.~Musatov and A.~V.~Radyushkin, Phys.\ Rev.\ D {\bf 61}, 074027 (2000).
